---
output:
  pdf_document:
    fig_height: 3
    fig_width: 5
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---

### STAT 230 - Analysis Component - Group 1B

```{r, include=FALSE}
library(mosaic)
library(readr)
library(GGally)
```

Names: Charlotte Kellogg, Anna Zhou, Alison Ortiz Dimas

(Tentative) Project Title: Predicting Housing Prices in King County, Washington State

Our project is about predicting housing prices in King County, WA, using ten quantitative variables. We are researching all houses that were sold in King County between May 2014 and May 2015 using thousands of observations from 372 dates.

### Read in the data

```{r}
# Must start with a data-read in command from Prof. Wagaman
theData <- read_csv("https://awagaman.people.amherst.edu/stat495/kc_house_data.csv")
theData <- mutate(theData, renovated = ifelse(yr_renovated>0,"Yes","No"))
```

### Summary command on data set

```{r}
glimpse(theData)

### Summary command on data set
````

```{r}
theMod <- lm(price ~ bedrooms+bathrooms+sqft_living+sqft_lot+floors+condition+yr_built+renovated, data = theData)
msummary(theMod)
vif(theMod)
```


```


### Data Codebook

List your variables and whether they are quantitative/qualitative (numeric vs categorical), along with other notes about the variables. Hint, to make a nice list, you need to put two spaces at the end of each line to force RMarkdown to start a new line. Check this out below:

Our variables are:  
Variable 1 -  date - categorical
Variable 2 -  bedrooms - numeric
Variable 3 -  bathrooms - numeric
Variable 4 - sqft_living - numeric
Variable 5 - sqft_lot - numeric
Variable 6 - floors - numeric
Variable 7 - condition - numeric
Variable 8 - sqft_basement - numeric
Variable 9 - yr_built - numeric
Variable 10 - renovated - categorical

### Analysis Plan

A rough outline for your proposed analysis, including univariate summaries, bivariate (or multivariate) relationships and plans for your model(s) and visualizations should follow via the sections below; as well as any additional thoughts about your randomization-based procedure. 

This section doesn't need to have plots for EVERY variable in your data set if you have many, but it needs to demonstrate that you've started exploring the data, identifying issues that arise, and are looking into what appropriate models might be. 

What you do here may be useful for your reports (i.e. may reuse it), so you may want to spend time making nice labels for plots, etc. 

#### Prelim Univariate Analysis

Obtain basic univariate descriptive statistics and graphs for variables relevant to your analysis.

```{r}
gf_dens(~bedrooms, data=theData)
gf_dens(~bathrooms, data=theData)
gf_dens(~sqft_living, data=theData)
gf_dens(~sqft_lot, data=theData)
gf_dens(~floors, data=theData)
gf_dens(~condition, data=theData)
gf_dens(~sqft_basement, data=theData)
gf_dens(~yr_built, data=theData)
tally(~renovated, data=theData)
gf_dens(~price, data = theData)

msummary(theData$price)
msummary(theData$bedrooms)
msummary(theData$bathrooms)
msummary(theData$sqft_living)
msummary(theData$sqft_lot)
msummary(theData$floors)
msummary(theData$condition)
msummary(theData$sqft_basement)
msummary(theData$yr_built)
msummary(theData$renovated)

```

COMMENT on what you see! Suggest doing this as you go. I.E. Do a summary for variable 1, then variable 2, etc. 
  
The density plots of bathrooms, bedrooms, sqft_living, sqft_lot, sqft_basement, and price are all highly skewed to the right, suggesting that the means are greater than the medians for those variables. There is a lot of variability within the yr_built variable. There are many more houses that were not renovated than there are that were.

It is very important to do this for your response variable(s), so be sure those are included here.

#### Prelim Multivariate Analysis

Scatterplots and side-by-side boxplots to examine bivariate relationships that will be useful for building your models. I.E. relationships between your response(s) and predictors. You can investigate relationships between predictors of interest as desired. 

```{r}

theData2 <- theData %>%  #selecting the variables that we decided to work with according to proposal
  select(c(price
           , sqft_above
           , sqft_basement
           , bedrooms
           , bathrooms
           , sqft_living
           , sqft_lot
           , condition
           , floors
           , yr_built
           , yr_renovated)) %>%
  filter(yr_renovated != 0) #removing values where there was no renovation

theData2$condition <- factor(theData2$condition) #converting condition into a categorical variable

gf_boxplot(price ~ condition, data = theData2)

ggplot(theData2, aes(sqft_above, price)) + geom_point() + geom_smooth(se = F) 
ggplot(theData2, aes(sqft_above, price, color=yr_built)) + geom_point() + geom_smooth(se = F) 
ggplot(theData2, aes(sqft_above, price, color=condition)) + geom_point() + geom_smooth(se = F) 

ggplot(theData2, aes(sqft_basement, price)) + geom_point() + geom_smooth(se = F) 
ggplot(theData2, aes(sqft_basement, price, color = condition )) + geom_point() + geom_smooth(se = F) 
ggplot(theData2, aes(sqft_basement, price, color = bathrooms )) + geom_point() + geom_smooth(se = F) 

ggplot(theData2, aes(bedrooms, price)) + geom_point() + geom_smooth(se=F) 
ggplot(theData2, aes(bedrooms, price, color = condition)) + geom_point() + geom_smooth(se=F) 

ggplot(theData2, aes(bathrooms, price)) + geom_point() + geom_smooth(se=F) 
ggplot(theData2, aes(bathrooms, price, color = condition)) + geom_point() + geom_smooth(se=F) 
ggplot(theData2, aes(bathrooms, price, color = bedrooms)) + geom_point() + geom_smooth(se=F) 

ggplot(theData2, aes(sqft_lot, price)) + geom_point() + geom_smooth(se = F) 
ggplot(theData2, aes(sqft_lot, price, color = bathrooms)) + geom_point() + geom_smooth(se=F) 
ggplot(theData2, aes(sqft_lot, price, color = condition)) + geom_point() + geom_smooth(se=F) 
ggplot(theData2, aes(sqft_lot, price, color = floors)) + geom_point() + geom_smooth(se=F) 

ggplot(theData2, aes(floors, price)) + geom_point() + geom_smooth(se = F) 
ggplot(theData2, aes(floors, price, color = bedrooms)) + geom_point() + geom_smooth(se = F) 
ggplot(theData2, aes(floors, price, color = sqft_lot)) + geom_point() + geom_smooth(se = F)


ggplot(theData2, aes(yr_built, price)) + geom_point() +  geom_smooth(se = F)
ggplot(theData2, aes(yr_built, price, color = condition)) + geom_point() +  geom_smooth(se = F)


ggplot(theData2, aes(yr_renovated, price)) + geom_point() + geom_smooth(se=F) 
ggplot(theData2, aes(yr_renovated, price, color=condition)) + geom_point() + geom_smooth(se=F) 



#gf_point or gf_boxplot #or if trying to do many at once
```

COMMENT on what you see! 

You can do scatterplots by color, etc. to do more than bivariate relationships. Other options include faceting to include more variables. 

We definitely expected there to be stronger relationships between price and condition, year built, numbers of bedrooms, and the total square feet of the lot. There are some clear outliers in the plots between price and sqft_above and in the plot between price and sqft_basement which might be making the relationship appear stronger than it actually is. We were particularly surprised to see that the relationship between sqft_lot and price wasn't as strong as we had thought it would have been. Again, we would have to examine the impact that condition had on the high selling price of certain homes despite a lower total lot area. We also need to examine the data and figure out whether all the data points were actually homes since there was one data point that had 33 bedrooms (which leads us to believe that it might be an apartment complex).


#### Randomization-Based Procedure Thoughts

One potential randomization-based procedure we could try to use is a randomization F-test. We can try multiple different forms of randomization F-tests based on what our earlier analysis indicates and depending on whether we choose an additive or interaction model. We can test for the main effects of each individual predictor by shuffling them, or also test for the overall model by shuffling the response. If we do end up choosing an interaction model, we can test for just the interaction effect while leaving the main effects intact by fitting the model without interaction and then permuting the residuals before adding them back to the fitted values. This allows us to see how the F-statistic of the original fit compares to the permuted stats. 

Thoughts on what procedure you might like to try; aim to use, etc. 

### Questions for me

Any questions you have for me.
